---
title: "Week10_Thoracic_Surgery"
author: "Akila Selvaraj"
date: "2/19/2022"
output:
  pdf_document: default
  html_document: default
---

```{r include=FALSE}
#setwd("G:/Users/a162940/Akila/Work/R/Projects/dsc520-master")
setwd("C:/Users/akila/Documents/Data Science/R_projects/dsc520-master")



library('foreign')
library("rio")

library(readxl)

Thoracic_df <- data.frame(read.arff("data/ThoraricSurgery.arff"))

```

# Logistic regression
```{r}

library('mlogit')

mlChat <- mlogit.data(Thoracic_df, choice = "Risk1Yr", shape = "wide")

Thoracic_glm <- glm(Risk1Yr ~ AGE + DGN + PRE4 + PRE5  + PRE6 + PRE7 + PRE8 + PRE9 +
                      PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 +
                      PRE30 + PRE32 , data = Thoracic_df, family = binomial(link='logit'))
```

# Summary of logistic model
```{r}
summary(Thoracic_glm)
```
Out of all the variables, few of the variables like PRE9T,PRE14OC14,PRE17T,PRE30T turned out to be significant.

PRE9T - The beta coefficient against this variable is 1.368. Let us convert this value into odds by taking the exp(1.368) = 3.927. The value indicates that the odds of an individual being in Risk1Yr increases by 3.9% than the one with PRE9 F value.

Out of 2 levels, F level became the reference, and thus all PRE variables are inferred in comparison to the referenced variable. That is how we interpret the categorical variables.

Age - The beta coefficient of the age variable is 0.009506, which is in the logit of odds terms. 
When we convert this to odds by taking exp(0.023362) we get 1.0095. 
The value indicates that as age increase by one more unit, then the odds of an individual being in the Risk1Yr F group will increase by 2%.

# Predicting outcome variable
```{r}
predicted_df <- predict(Thoracic_glm, Thoracic_df, type = "response")
summary(predicted_df)
```

# Percentage of correct predictions - Accuracy of the model
```{r}
library("performance")
performance_pcp(Thoracic_glm, ci = 0.95, method = "Herron", verbose = TRUE)
```
AS PCP is above 50%, we can say that the PCP of the full model is considerably above the null model's PCP.

The likelihood-ratio test indicates that the model has a significantly better fit than the null-model as p is less than 0.05.

The accuracy of the model from percentage of correct predictions is 77%.
